---
title: "Ngiam and Lee (2026) - perceptualReproduction"
author: "William XQ Ngiam"
date: "`r Sys.Date()`"
output: html_document
---

## Overview

This file is associated with the preprint titled ["Similarity judgments and visual working memory do not share the same cognitive representation"](https://osf.io/preprints/psyarxiv/fm9vz_v1) by William XQ Ngiam (Adelaide University) and Michael D Lee (University of California Irvine. This research project models the underlying psychological space in perceptual tasks, working memory tasks, and similarity judgment tasks.

The following model recovers the underlying psychological space for the orientation **perceptual reproduction** task in [Tomic and Bays (2024)](https://psycnet.apa.org/doi/10.1037/xlm0001172). In this task, a line at an angle (0 to 180 degrees) is shown to the subject and remains on the screen. The subject reproduces the orientation as closely as possible.

The following code builds a Bayesian Thurstonian model using JAGS. The underlying Thurstonian model is an attempt to recover the representation (much like multidimensional scaling) of the orientations. Then, this model could be used to predict performance on other tasks; such as a working memory task with these orientations.

### Requirements

The following files should be saved in your working directory:

* The data file ("oriPercData_clean.csv") - [downloadable here](https://github.com/WilliamNgiam/orientationModeling/blob/main/data/oriPercData_clean.csv).
* The model text file ("perceptualReproduction_jags.txt") - [downloadable here](https://github.com/WilliamNgiam/orientationModeling/blob/main/perceptualReproduction/perceptualReproduction_jags.txt).

You will also need to have installed JAGS on your device. The installer can be accessed here: [https://sourceforge.net/projects/mcmc-jags/files/]. The following code was built using JAGS 4.3.2.

### Setup

```{r setup}
# Load required packages
library(R2jags) # v0.8-9
library(tidyverse) # v2.0.0
library(coda) # v0.19-4.1
library(ggmcmc) # v1.5.1.2

# Provide R session output 
sessionInfo()

# Set RNG seed for reproducibility
set.seed(123)
```

### Load data

```{r read data, warning = FALSE, message = FALSE}
# Load in Tomic and Bays (2024) dataset
percData <- read_csv(paste0(getwd(),"/oriPercData_clean.csv"))

# Create intervals to match the Tomic and Bays (2024) perceptual task.
binIntervals = seq(0,pi,length.out = 73)

# Assign indices to the closest interval for the target (stimulus) values and response values
percData <- percData %>% 
  mutate(response = round(response,digits = 5),
         target = round(target,digits = 5)) %>% 
  # Make all values positive
  mutate(response = response/2,
         target = target/2) %>% 
  mutate(response = case_when(response < 0 ~ response + pi,
                              .default = response),
         target = case_when(target < 0 ~ target + pi,
                            .default = target)) %>%
  # Adjust response values to account for circular angle
  mutate(response = case_when(abs(response-target) > abs(response-target-pi) ~ response-pi,
                              .default = response)) %>% 
  mutate(response = case_when(abs(target-response) > abs(target-response-pi) ~ response+pi,
                              .default = response)) %>% 
  rowwise() %>% 
  # Match target values to the intervals on the semi-circle
  mutate(targetIndex = which.min(abs(target-binIntervals))) %>% 
  # The target values closest to pi are re-indexed to the first bin (0 degrees)
  mutate(targetIndex = case_when(targetIndex == 73 ~ 1,
                                 .default = targetIndex))

# Set up the true values to be recovered
muTruth = binIntervals[-73] # Remove the last index because that is the point at which the circle wraps around
sigmaTruth = 0.01

# Assign data to vectors
s <- as.integer(percData$targetIndex)
y <- as.numeric(percData$response)
nTrials <- as.integer(length(percData$response))
nStimuli <- as.integer(length(muTruth))
```

### Read in data for JAGS

```{r read in data}
data.jags <- list("s","y","nTrials","nStimuli")
bayes.mod.params <- c("mu","sigma")
bayes.mod.inits <- function() {
  list("sigma" = runif(1)*pi)
}
```

### Run model fit

```{r run model fit}
bayes.mod.fit <- jags(data = data.jags,
                      inits = bayes.mod.inits,
                      parameters.to.save = bayes.mod.params,
                      n.chains = 8,
                      n.iter = 8000,
                      n.burnin = 2000,
                      model.file = paste0(getwd(),"/perceptualReproduction_jags.txt"))

```
### Evaluate model

```{r evaluate model}
print(bayes.mod.fit)
```

```{r evaluate model as mcmc}
bayes.mod.fit.mcmc <- as.mcmc(bayes.mod.fit)
summary(bayes.mod.fit.mcmc)
```

### Plots

```{r plot model}
# Convert into ggmcmc object
bayes.mod.fit.gg <- ggs(bayes.mod.fit.mcmc)

ggs_caterpillar(bayes.mod.fit.gg,
                thick_ci = c(0.45,0.55),
                family = "mu",
                sort = FALSE)

ggs_density(bayes.mod.fit.gg,
            family = "sigma")
```